---
title: "Machine Learning Project"
output:
  html_document:
    df_print: paged
---





```{r warning = FALSE, message = FALSE}
# Suppress dplyr summarise grouping warning messages
options(dplyr.summarise.inform = FALSE)

## Add R libraries here
library(tidyverse)
library(tidymodels)
library(dplyr)
library(ggplot2)

# Load the dataset
telecom_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/telecom_df.rds'))

```



# Data Analysis

# Question 1
**Question**: 
How is customer's number of months of joining the company related to their canceled service?

**Answer**:
It is observed that majority of the customers tend to cancel their services within 5 months of their joining. 159 customers have cancelled their services with first 5 months. The second highest period is within 5 to 10 months, where 58 customers have cancelled their services.

Many Customers tend to cancel their services in  0 to 15 months of their service.


```{r}
q1 = telecom_df%>% mutate(cust_suration = cut_width(months_with_company, width= 5, boundary = 0))


q1%>% filter(canceled_service == 'yes') %>% group_by(cust_suration) %>% summarize(
  number_of_cust = n(),
  min_monthly_charges = min(monthly_charges),
  avg_monthly_charges = mean(monthly_charges),
  median_monthly_charges = median(monthly_charges),
  max_monthly_charges = max(monthly_charges)
)



```
```{r}
ggplot(telecom_df, aes(months_with_company, fill = canceled_service)) + 
  geom_histogram( color = 'white' , bins= 15) + facet_wrap(~canceled_service ,nrow = 2) +
  labs(title = 'Customers duration with company in months affecting their canceled service' , x = 'Months with company', y = 'Count')
```




# Question 2
**Question**:
Is there any relation between  bill payment mode and canceled services?

**Answer**:

Electronic Check mode of payment has seen the highest number of customers cancelling their service. It is seen that 256 people accounting for 61.8 % of the canceled service customers pay their bills through electronic check.

The  payment mode facing second highest number of customers cancelling their services is bank draft. 57 customers accounting 13.34% of the canceled service customers pay their bills through bank draft.

The average monthly charges paid by electronic check that is $83.96228 and bank draft that is $85.14486 are highest compared to other payment modes

```{r}

telecom_df %>%  group_by(payment_method) %>% summarize(
  count = n(),
  avg_monthly_charges = mean(monthly_charges),
  median_monthly_charges = median(monthly_charges)
)


```
```{r}
telecom_df %>%  group_by(payment_method,canceled_service) %>% summarize(
  count = n(),
  avg_monthly_charges = mean(monthly_charges),
  median_monthly_charges = median(monthly_charges)
)

```

```{r}
telecom_df %>% filter(canceled_service== 'yes') %>% group_by(payment_method) %>% summarize(
  count = n(),
  avg_monthly_charges = mean(monthly_charges),
  percent = (count/427) * 100

)
```

```{r}
ggplot(telecom_df, aes(x = payment_method,fill = canceled_service)) + geom_bar(stat = "count") +
  facet_wrap(~canceled_service, nrow = 2)
```


# Question 3
**Question**:
Which cellular service is experiencing more number of customers cancelling their service?

**Answer**:
Many Customers with multiple lines cellular service are opting out their services. It is seen that 223 multiple lines customers have opted out the service. It is also observed that customers using multiple line cellular service and their monthly expenses in the range of $70 to $100 are more likely to discontinue their services.

```{r}
telecom_df %>% group_by(canceled_service,cellular_service) %>% summarize(
  count = n()
)

```

```{r}
q3 = telecom_df%>% mutate(monthly_charges_range = cut_width(monthly_charges, width= 10, boundary = 0))

q3 %>% filter(canceled_service == 'yes' & cellular_service =='multiple_lines') %>% 
  group_by(monthly_charges_range) %>%
  summarize(
  count = n(),
  avg_monthly_bill = mean(monthly_charges)
)
```

```{r}
ggplot(telecom_df, aes(monthly_charges, fill = canceled_service)) + 
  geom_histogram( color = 'white' , bins= 15) + facet_grid(cellular_service ~ canceled_service ) +
  labs(title = 'Customers monthly charges' , x = 'Monthly charges', y = 'Count')
```



# Question 4
**Question**:
Which contract customers tend to have higher avg_call_mins and their monthly bill?

**Answer**:
Customers opting for month to month contract have higher cancellations in the service. 90.8% percent of the canceled services are from month to month contract. The average call in minutes and the average monthly charges of the customers cancelling the services and belong to month to month contract is 353 mins and $79 which is comparatively less than other contract customers.
The offers offered by the company is beneficial for customers with higher average call in minutes

```{r}
telecom_df %>% group_by(contract) %>% summarize(
  
  count = n(),
  avg_call_mins = mean(avg_call_mins),
  avg_monthly_charges = mean(monthly_charges)
)
```

```{r}
telecom_df %>% filter(canceled_service == 'yes')%>% group_by(contract,canceled_service) %>% summarize(
  
  count = n(),
  avg_call_mins = mean(avg_call_mins),
  avg_monthly_charges = mean(monthly_charges),
  percetange = (count/427) * 100
)
```

```{r}
ggplot(telecom_df,aes(x = contract, y = avg_call_mins, fill = canceled_service)) +
  geom_boxplot() 
```



# Question 5
**Question**:
Is there any relation between canceled_service and Average international calls in mins?

**Answer**:

It can be observed that cancelled service customers tend to have less average international mins. The average international call in mins for the cancelled service customers is 93.5 mins where as the average internation call mins for the cancelled service customers is 113.45 mins which is high.

Therefore, customers with higher usage for international call (more than 100 mins in average) tend to not cancel their service.

```{r}
telecom_df %>% group_by(canceled_service) %>% summarize(
  count = n(),
  avg_intl_mins = mean(avg_intl_mins),
  median_intl_mins = median(avg_intl_mins)
)
```



```{r}
library(plotly)
fig = plot_ly(
  data = telecom_df,
  x = ~canceled_service,
  y = ~avg_intl_mins,
  type = "violin",
  color = ~canceled_service,
  side = "positive",
  meanline = list(visible = T),
  colors = c("red", "blue")
) %>% layout(title = 'Relation between Average International calls and cancled service')
fig

```





# Machine Learning

# Model 1
# Decision tree 


```{r}
library(tidymodels)
library(vip)
library(rpart.plot)
set.seed(271)

telecom_split <- initial_split(telecom_df, prop = 0.75,
                               strata = canceled_service)

telecom_training <- telecom_split %>% 
  training()

telecom_test <- telecom_split %>%
  testing()

telecom_folds <- vfold_cv(telecom_training, v = 5)

telegram_recipe <- recipe(canceled_service ~ ., data = telecom_training) %>% 
  step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
  step_normalize(all_numeric(), -all_outcomes()) %>% 
  step_dummy(all_nominal(), -all_outcomes())

telegram_recipe %>% 
  prep(training = telecom_training) %>% 
  bake(new_data = NULL)

tree_model <- decision_tree(cost_complexity = tune(),
                            tree_depth = tune(),
                            min_n = tune()) %>% 
  set_engine('rpart') %>% 
  set_mode('classification')

tree_workflow <- workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(telegram_recipe)

tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          min_n(), 
                          levels = 2)

tree_grid

tree_grid <- grid_regular(parameters(tree_model), 
                          levels = 2)

set.seed(314)

tree_tuning <- tree_workflow %>% 
  tune_grid(resamples =  telecom_folds,
            grid = tree_grid)

best_tree <- tree_tuning %>% 
  select_best(metric = 'roc_auc')

# View the best tree parameters
best_tree


final_tree_workflow <- tree_workflow %>% 
  finalize_workflow(best_tree)
```

```{r}
#Training data
tree_wf_fit <- final_tree_workflow %>% 
  fit(data = telecom_training)


tree_fit <- tree_wf_fit %>% 
  extract_fit_parsnip()


vip(tree_fit)

```


```{r}
rpart.plot(tree_fit$fit, roundint = FALSE, extra = 2)

rf_last_fit <- final_tree_workflow %>% 
  last_fit(telecom_split)


```

```{r}
tree_last_fit <- final_tree_workflow %>% 
                 last_fit(telecom_split)
tree_last_fit %>% collect_metrics()
```



```{r}

tree_last_fit %>% collect_predictions() %>% 
                  roc_curve(truth  = canceled_service, estimate = .pred_yes) %>% 
                  autoplot()
```




```{r}
tree_predictions <- tree_last_fit %>% collect_predictions()

conf_mat(tree_predictions, truth = canceled_service, estimate = .pred_class)
```


# Model 2
# LDA
```{r}
library(tidymodels)
library(discrim)
library(rlang)
library(klaR)
library(labelled)
lda_model <- discrim_regularized(frac_common_cov = 1) %>% 
  set_engine('klaR') %>% 
  set_mode('classification')

telecom_recipe <- recipe(canceled_service ~ ., data = telecom_training) %>% 
  step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
  step_normalize(all_numeric(), -all_outcomes()) %>% 
  step_dummy(all_nominal(), -all_outcomes())

telecom_recipe %>% 
  prep(training = telecom_training) %>% 
  bake(new_data = NULL)

lda_wf <- workflow() %>% 
  add_model(lda_model) %>% 
  add_recipe(telecom_recipe)

last_fit_lda <- lda_wf %>% 
  last_fit(split = telecom_split)

last_fit_lda %>% collect_metrics()



```


```{r}
lda_predictions <- last_fit_lda %>% 
                     collect_predictions()

lda_predictions


lda_predictions %>% 
  roc_curve(truth = canceled_service, .pred_yes) %>% 
  autoplot()

```
```{r}

conf_mat(lda_predictions, truth = canceled_service, estimate = .pred_class)


```


# Model 3
# KNN
```{r}
knn_model <- nearest_neighbor(neighbors = tune()) %>% 
  set_engine('kknn') %>% 
  set_mode('classification')
knn_wf <- workflow() %>% 
  add_model(knn_model) %>% 
  add_recipe(telecom_recipe)
k_grid <- tibble(neighbors = c(10, 20, 30, 50, 75, 100, 125, 150))
k_grid
knn_tuning <- knn_wf %>% 
  tune_grid(resamples = telecom_folds,
            grid = k_grid)
knn_tuning %>% show_best('roc_auc')

best_k <- knn_tuning %>% 
  select_best(metric = 'roc_auc')

## View model
best_k

final_knn_wf <- knn_wf %>% 
  finalize_workflow(best_k)

last_fit_knn <- final_knn_wf %>% 
  last_fit(split = telecom_split)

knn_predictions <- last_fit_knn %>% 
  collect_predictions()


```
```{r}
knn_predictions %>% 
  roc_curve(truth = canceled_service, estimate = .pred_yes) %>% 
  autoplot()

```
```{r}
last_fit_knn %>% collect_metrics()
```




```{r}
conf_mat(knn_predictions, truth = canceled_service, estimate = .pred_class)

```


# Model 4
# Random Forest

```{r}
library(mod)
library(base)
library(ranger)
rf_model <- rand_forest(mtry = tune(),
                        trees = tune(),
                        min_n = tune()) %>% 
            set_engine('ranger', importance = "impurity") %>% 
            set_mode('classification')
rf_workflow <- workflow() %>% 
               add_model(rf_model) %>% 
               add_recipe(telecom_recipe)
set.seed(314)

rf_grid <- grid_random(mtry() %>% range_set(c(2, 4)),
                       trees(),
                       min_n(),
                       size = 10)
rf_tuning <- rf_workflow %>% 
             tune_grid(resamples = telecom_folds,
                       grid = rf_grid)
rf_tuning %>% show_best('roc_auc')
best_rf <- rf_tuning %>% 
           select_best(metric = 'roc_auc')
best_rf
final_rf_workflow <- rf_workflow %>% 
                     finalize_workflow(best_rf)
rf_wf_fit <- final_rf_workflow %>% 
             fit(data = telecom_training)
rf_fit <- rf_wf_fit %>% 
          extract_fit_parsnip()
```

```{r}
vip(rf_fit)
```

```{r}
rf_last_fit <- final_rf_workflow %>% 
               last_fit(telecom_split)
rf_last_fit %>% collect_metrics()
```

```{r}

rf_last_fit %>% collect_predictions() %>% 
                roc_curve(truth  = canceled_service, estimate = .pred_yes) %>% 
                autoplot()
```
```{r}
rf_predictions <- rf_last_fit %>% collect_predictions()

conf_mat(rf_predictions, truth = canceled_service, estimate = .pred_class)
```


# Model 5
# QDA

```{r}
qda_model <- discrim_regularized(frac_common_cov = 0) %>% 
             set_engine('klaR') %>% 
             set_mode('classification')
qda_wf <- workflow() %>% 
          add_model(qda_model) %>% 
          add_recipe(telecom_recipe)

last_fit_qda <- qda_wf %>% 
                last_fit(split = telecom_split)

```
```{r}
last_fit_qda %>% collect_metrics()
```


# Model 6
# Logistic Regression
```{r}
logistic_model <- logistic_reg() %>% 
                  set_engine('glm') %>% 
                  set_mode('classification')

telegram_recipe <- recipe(canceled_service ~ ., data = telecom_training) %>% 
  step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
  step_normalize(all_numeric(), -all_outcomes()) %>% 
  step_dummy(all_nominal(), -all_outcomes())

telecom_wf <- workflow() %>% 
            add_model(logistic_model) %>% 
            add_recipe(telecom_recipe)

telecom_logistic_fit <- telecom_wf %>% 
                      fit(data = telecom_training)

telecom_trained_model <- telecom_logistic_fit %>% 
                       extract_fit_parsnip()

```

```{r}
vip(telecom_trained_model)
```

```{r}
predictions_categories <- predict(telecom_logistic_fit, 
                                  new_data = telecom_test)

predictions_categories
```
```{r}
predictions_probabilities <- predict(telecom_logistic_fit, 
                                     new_data = telecom_test, 
                                     type = 'prob')

predictions_probabilities

```
```{r}


test_results <- telecom_test  %>%
                bind_cols(predictions_categories) %>% 
                bind_cols(predictions_probabilities)

test_results
```


```{r}
roc_curve(test_results, 
          truth = canceled_service,
          estimate = .pred_yes)
```
```{r}
roc_curve(test_results, 
          truth = canceled_service, 
          estimate = .pred_yes) %>% 
  autoplot()
```

```{r}
last_fit_logistic <- telecom_wf %>% 
                last_fit(split = telecom_split)

last_fit_logistic %>% collect_metrics()
```
```{r}
conf_mat(test_results, 
         truth = canceled_service, 
         estimate = .pred_class)
```


